Data migration under a Changing Schema

### 2022-12-27 2022 Migration/articleMigration.tex
refined the migration steps.

   To migrate system $\infsys$ to $\infsys'$, we take the following steps:
\begin{enumerate}
   \item Let the existing system $\infsys$ have state $\dataset$ and schema $\schema$.
         A developer defines a desired system $\infsys'$ by specifying an initial dataset $\dataset'$, a schema $\schema'$, and some functionality\footnote{For now, we ignore the functionality.}.
         Schema $\schema'$ is the schema that governs the system after the migration.
         After the migration, the system must contain the triples from both $\dataset$ and $\dataset'$, to preserve the existing data.
         If the developer wants to add some new data into the system on the moment of transition (MoT),
         she specifies that in $\dataset'$.
         That might be necessary to initialize specific new features.
         If $\schema$ equals $\schema'$, we call the migration trivial because a migration is not necessary.
   \item The generator generates a migration script $\migrsys$ based on $\schema$ and $\infsys'$
         to migrate the data and define the semantics during the migration phase as far as a generator can reasonably predict.
         The schema $\id{migrationrules}$ represents the enforce rules that migrate the existing data to the desired dataset.
   \item A migration engineer changes $\id{migrationrules}$ to accommodate the non-standard requirements of the migration,
         yielding migration script $\migrsys'$.
         This schema also contains data structures and rules that are needed for the duration of the migration phase.
         The migration engineer does not alter the existing system $\infsys$ nor the specification of the desired system $\infsys'$ that the developer has produced.
         As a result, the migration engineer can always restore the existing system because it is part of the migration system.
   \item The production environment switches instantly from $\infsys$ to $\migrsys'$,
         while the data of the existing system remains intact.
         This can be realized by configuring the ingress system to open the route to the desired system,
         while keeping the existing system accessible.
         This event marks the MoT.
         From this point onward, both the existing functionality and the desired functionality are available to users.
   \item The migration system $\dataset_{\migrsys'}$ must be at rest before the MoN,
         to ensure that the migration has completed before the migration becomes irreversible.
   \item The business has the opportunity to evaluate the change, which must end in a go/no-go decision.
   \item In case of a go-decision, the migration engineer can disable the undo-facility as soon as the migration system is at rest.
         In case of a no-go, the migration engineer restores the existing system.
         This event marks the moment of no return (MoN)
   \item Users may have to perform manual migration steps.
         These are specified as process rules in $\id{migrationrules}$.
         % Dit zijn "kurk" regels.
         The system ensures that the amount of work needed to perform these steps is finite.
   \item The patron of the system can decide whether to carry on with the migration or to roll back to the existing system.
         She can do that incrementally by switching off the interfaces one-by-one.
   \item The system retains the data, the schema, and the functionality that is covered by $\schema'$, dropping everything else.
\end{enumerate}

>>>>

   To migrate system $\infsys$ to $\infsys'$, we take the following steps:
\begin{enumerate}
   \item Let the existing system $\infsys$ have state $\dataset$ and schema $\schema$.
         A developer defines a desired system $\infsys'$, which contains a dataset of its own, $\dataset'$, a schema $\schema'$, and some functionality\footnote{For now, we ignore the functionality.} as illustrated in figure~\ref{fig:post-migration}.
         After the migration, the system will contain the triples from $\dataset$, to preserve the existing data.
         The developer may want to add some new data into the system,
         for instance to initialize new features on the moment of transition (MoT).
         She specifies such data in $\dataset'$.
         If $\schema$ equals $\schema'$, we call the migration trivial because a migration is not necessary.
   \item The generator generates a migration script $\migrsys$ based on $\schema$ and $\infsys'$
         to migrate the data and define the semantics during the migration phase as far as a generator can reasonably predict.
         The schema $\id{migrationrules}$ represents the enforce rules that migrate the existing data to the desired dataset.
   \item A migration engineer changes $\id{migrationrules}$ to accommodate the non-standard requirements of the migration,
         yielding migration script $\migrsys'$.
         This schema also contains data structures and rules that are needed for the duration of the migration phase.
         The migration engineer does not alter the existing system $\infsys$ nor the specification of the desired system $\infsys'$ that the developer has produced.
         As a result, the migration engineer can always restore the existing system because it is part of the migration system.
   \item The production environment switches instantly from $\infsys$ to $\migrsys'$,
         while the data of the existing system remains intact.
         This can be realized by configuring the ingress system to open the route to the desired system,
         while keeping the existing system accessible.
         This event marks the MoT.
         From this point onward, both the existing functionality and the desired functionality are available to users as illustrated in figure~\ref{fig:migration phase}.
   \item To ensure that all data is migrated before the the migration engineer makes the migration irreversible,
         we require that the rules of the the migration system $\rules_{\migrsys'}$ are at rest before the migration engineer takes down the existing system.
         Users (i.e.\ the business) must do the work to satisfy the manually enforced rules.
         Ampersand guarantees that enforce rules will eventually be satisfied.
   \item To give the business an opportunity to evaluate the change,
         we can require a go/no-go decision before the the migration engineer makes the migration irreversible.
   \item In case of a go-decision, the migration engineer can disable the undo-facility as soon as the migration system is at rest.
         In case of a no-go, the migration engineer restores the existing system.
         This event marks the moment of no return (MoN).
   \item The system retains the data, the schema, and the functionality that is covered by $\schema'$, dropping everything else.
         This establishes the post-migration situation (figure~\ref{fig:post-migration}).
\end{enumerate}
